---
title: "Summary Description of Tiger and Leopard Image Data in the \n ExtractCompare Database"
author: "Chandan Kumar Pandey"
output:
  word_document: default
  fig_width: 10
  fig_height: 8
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---
## This report summarizes the image data of tigers and leopards in the ExtractCompare database of CWS.

### 1. Total Number of images of tigers and leopards
The EC database has a total of 43781 images belonging to 3367 individual tigers and leopards obtained largely through camera trap surveys from 12 Wildlife Reserves and opportunistic collection from a larger network of forested areas neighboring these camera trap “sites”. The opportunistic collection included images of dead tigers, seized pelts, photographs contributed by naturalists and others. Camera trap images contribute to nearly 93% of 25,037 tiger images and 99% of 18,739 leopard images. These images belong to a total of 1984 individual tigers and 1654 individual leopards. Camera trap surveys contribute to nearly 73% and 92% of individual tigers and leopards respectively.

```{r setup, include=FALSE}
# Load configuration
if(file.exists("config.R")) {
  source("config.R")
} else {
  # Fallback defaults if config.R is missing in knitting environment
  TIGER_DATA_FILE <- "Animal_cature_details.csv"
  LEOPARD_DATA_FILE <- "leopard_capture_details.csv"
}
```

```{r loading the data in the file,warning=FALSE,echo=FALSE,comment=FALSE,message=FALSE}
if (file.exists(TIGER_DATA_FILE)) {
  Tiger<-read.csv(TIGER_DATA_FILE, header = T)
  Tiger$name<-rep("Tiger",nrow(Tiger))
  Tiger<-Tiger[!duplicated(Tiger),]
} else {
  stop(paste("Tiger data file not found:", TIGER_DATA_FILE))
}

if (file.exists(LEOPARD_DATA_FILE)) {
  leopard<-read.csv(LEOPARD_DATA_FILE, header = T)
  leopard$name<-rep("leopard",nrow(leopard))
  leopard<-leopard[!duplicated(leopard),]
} else {
  stop(paste("Leopard data file not found:", LEOPARD_DATA_FILE))
}

require(dplyr)
require(ggplot2)

# Ensure data has expected columns before binding
required_columns <- c("id", "CPMD", "name", "side", "time_seen", "site")
missing_cols_tiger <- setdiff(required_columns, names(Tiger))
missing_cols_leopard <- setdiff(required_columns, names(leopard))

if(length(missing_cols_tiger) > 0) {
  warning(paste("Tiger data missing columns:", paste(missing_cols_tiger, collapse=", ")))
}
if(length(missing_cols_leopard) > 0) {
  warning(paste("Leopard data missing columns:", paste(missing_cols_leopard, collapse=", ")))
}

animal_data<-bind_rows(Tiger,leopard)
```

### 2. Number of images per individual tiger and leopard

```{r CPMD, warning=F,echo=F,message=F,fig.width=10, fig.height=8}
require(ggplot2)
require(gridExtra)
library(tidyr)
CPMD<-group_by(animal_data[is.na(animal_data$CPMD)==F,],CPMD,name)%>%
  summarise(value=n(),indivi=n_distinct(id))%>%
  as.data.frame()
CPMD$CPMD<-ordered(CPMD$CPMD,levels=c("CT","PH","CP","DE","FD","PE","MV"))
spread_CPMD<-spread(CPMD[,-4],key = c("CPMD"),value = c("value"),fill = 0)
spread_CPMD$rest_all<-rowSums(spread_CPMD[,4:8])
spread_CPMD<-spread_CPMD[,c(1,2,3,9)]
CPMD_img<-spread_CPMD%>%pivot_longer(!name, names_to="CPMD",values_to="count")%>%as.data.frame()
CPMD_img$name<-ordered(CPMD_img$name,levels=c("Tiger","leopard"))
cmpd_plot<-ggplot(data = CPMD_img,aes(y=count,x=CPMD,fill=as.factor(name)))
number<-cmpd_plot+geom_bar(stat="identity",position =position_dodge(0.9) )+
  geom_text(aes(label=count),hjust=0.25,vjust=-1,size=3,position = position_dodge(0.9),angle=40)+
  ylim(c(0,28000))+
  labs(x="",y="Number of images",label="Species")+
  theme_classic()+
  scale_fill_manual(name="",values=c("Darkblue","green"),labels=c("Tiger","Leopard"))+
  scale_x_discrete(labels=c("Camera Trap Photos","Photos by Naturalists","Other Photos"))

#indviduals
spread_ind<-spread(CPMD[,-3],key = c("CPMD"),value = c("indivi"),fill = 0)
spread_ind$rest_all<-rowSums(spread_ind[,4:8])
spread_ind<-spread_ind[,c(1,2,3,9)]
CPMD_ind<-spread_ind%>%pivot_longer(!name, names_to="CPMD",values_to="count")%>%as.data.frame()
CPMD_ind$name<-ordered(CPMD_ind$name,levels=c("Tiger","leopard"))
cmpd_plot_ind<-ggplot(data = CPMD_ind,aes(y=count,x=CPMD,fill=as.factor(name)))
ind<-cmpd_plot_ind+geom_bar(stat="identity",position =position_dodge(0.9) )+
  geom_text(aes(label=count),hjust=0.25,vjust=-1,size=3,position = position_dodge(0.9),angle=40)+
  ylim(c(0,2000))+
  labs(x="",y="Number of individuals",label="")+
  theme_classic()+
  scale_fill_manual(name="",values=c("Darkblue","green"),labels=c("Tiger","Leopard"))+
  scale_x_discrete(labels=c("Camera Trap Photos","Photos by Naturalists","Other Photos"))
grid.arrange(number,ind,nrow=2)
```

### 2. Number of images per individual tiger and leopard

It is clear that **`r round(spread_CPMD[1,2]/rowSums(spread_CPMD[1,-1]),2)*100`%** leopards and **`r round(spread_CPMD[2,2]/rowSums(spread_CPMD[2,-1]),2)*100`%** Tigers images are from camera trap.  Therefore we will only be using the image that were capture using camera traps. 

```{r historgrams of number of images per ind,echo=FALSE,warning=FALSE,message=FALSE,fig.width=10, fig.height=8}
library(ggplot2)
#tiger
camera_trap_data_tiger<-Tiger[Tiger$CPMD=="CT",]
camera_trap_data_tiger<-camera_trap_data_tiger[is.na(camera_trap_data_tiger$CPMD)==F,]
ploting_hist_tiger<-group_by(camera_trap_data_tiger,id)%>%summarise(count=n())%>%as.data.frame()
ploting_hist_tiger<-ploting_hist_tiger[order(ploting_hist_tiger$count),]
ploting_hist_tiger$intervel<-c("<=10","11-20","21-30","31-40", "41-50", ">50")[
  findInterval(ploting_hist_tiger$count,c(0,11,21,31,41,51,Inf))
]
ploting_hist_tiger$intervel<-ordered(ploting_hist_tiger$intervel,levels=c("<=10","11-20","21-30","31-40", "41-50", ">50"))
final_plot_tiger<-group_by(ploting_hist_tiger,intervel)%>%summarise(count=n())%>%as.data.frame()
#leopard
camera_trap_data_leo<-leopard[leopard$CPMD=="CT",]
camera_trap_data_leo<-camera_trap_data_leo[is.na(camera_trap_data_leo$CPMD)==F,]
ploting_hist_leo<-group_by(camera_trap_data_leo,id)%>%summarise(count=n())%>%as.data.frame()
ploting_hist_leo<-ploting_hist_leo[order(ploting_hist_leo$count),]
ploting_hist_leo$intervel<-c("<=10","11-20","21-30","31-40", "41-50", ">50")[
  findInterval(ploting_hist_leo$count,c(0,11,21,31,41,51,Inf))
]
ploting_hist_leo$intervel<-ordered(ploting_hist_leo$intervel,levels=c("<=10","11-20","21-30","31-40", "41-50", ">50"))
final_plot_leo<-group_by(ploting_hist_leo,intervel)%>%summarise(count=n())%>%as.data.frame()
final_plot_leo$name<-rep("Leopard",nrow(final_plot_leo))
final_plot_tiger$name<-rep("Tiger",nrow(final_plot_tiger))
final_plot<-bind_rows(final_plot_leo,final_plot_tiger)
final_plot$name<-ordered(final_plot$name,levels=c("Tiger","Leopard"))
hist_p<-ggplot(data = final_plot,aes(y=count,x=intervel,fill=name))
hist_p+geom_bar(stat="identity",position = position_dodge(0.9))+labs(x="Number of Images per Individual",y="Number of Individuals")+theme_classic()+
  ylim(c(0,1200))+geom_text(aes(label=count),hjust=0.2,vjust=-1,size=2,position = position_dodge(0.9),angle=20)+
  theme(axis.text.x = element_text(size = 8))+
  scale_fill_manual(name="",values=c("Darkblue","green"),labels=c("Tiger","Leopard"))
```

Nearly 65% and 73% of the camera-trapped individual tigers and individual leopards have less than or equal to 10 images per individual, while only 7% and 5% of individual tigers and leopards have more than 50 images per individual.

### 3. Number of camera trap images on different flanks

The number of right and left flank images per individual tiger and leopard are unequal, although nearly 90% of the individuals have almost equal number of right and left flank images. 


```{r flanks plot,echo=F,message=F,warning=F,fig.width=10, fig.height=8}
camera_trap_data<-bind_rows(camera_trap_data_leo,camera_trap_data_tiger)
camera_trap_data$side[grep("L",camera_trap_data$side)]<-"L"
camera_trap_data$side[grep("R",camera_trap_data$side)]<-"R"
flanks<-group_by(camera_trap_data,id,side,name)%>%summarise(number=n())%>%as.data.frame()

library(tidyr)
require(gridExtra)
flank_plot<-spread(data = flanks,key = "side",value = "number",fill = 0)
flank_plot<-flank_plot[,-5]
flank_plot<-flank_plot[is.na(flank_plot$name)==F,]
flank_plot$name<-stringr::str_replace(flank_plot$name,"leopard","Leopard")
flank_plot$name<-ordered(flank_plot$name,levels=c("Tiger","Leopard"))

spread_plot<-ggplot(data =  flank_plot,aes(x=log10(L),y=log10(R),color=name))
spread_plot+geom_point(shape=19,size=1.6)+labs(x="Number of left flank per individuals \n (in log scale base 10)",y="Number of right flank per individuals\n(in log base 10 scale)")+
  geom_abline(slope = 1, intercept = 0)+theme_classic()+scale_color_manual(values=c("Darkblue","green"),labels=c("Tiger","Leopard"),name="")

```

### 4.Day and Night-time images 



```{r traing with time,echo=F,warning=F,message=F,fig.width=10, fig.height=8}
require(lubridate)
require(data.table)
camera_trap_data<-camera_trap_data[is.na(camera_trap_data$time_seen)==F,]
camera_trap_data$time_seen<-as_datetime(camera_trap_data$time_seen)

time_data<-as_datetime(camera_trap_data$time_seen,tz="GMT")
time_data<-format(time_data,format = "%H%M%S")
time_data<-time_data%>%na.omit()
Time_block=as.POSIXct(strptime(c("060000","180000"),"%H%M%S"),tz="GMT")
Time_block<-format(Time_block,format = "%H%M%S")
time_of_day<-case_when(
between(time_data,Time_block[1],Time_block[2]) ~"morning",
        TRUE ~"night")
Time_data_frame<-data.frame(time=time_data,class=time_of_day,name=camera_trap_data$name)
Bar_chart<-group_by(Time_data_frame,class,name)%>%summarise(count=n())%>%as.data.frame()
Bar_chart$name<-stringr::str_replace(Bar_chart$name,"leopard","Leopard")
Bar_chart$name<-ordered(Bar_chart$name,levels=c("Tiger","Leopard"))
barp<-ggplot(data = Bar_chart,aes(x=class,y=count,fill=name))
barp+geom_bar(stat = "identity",color="black",position = position_dodge(0.9))+
  theme_classic()+labs(x="",y="Number of Images")+
  scale_x_discrete(breaks=unique(Bar_chart$class),labels=c("Day-time (0600 - 1800 Hrs)","Night-time (1801 - 0559 Hrs)"))+
  geom_text(aes(label=count),hjust=0.2,vjust=-1,size=5,position = position_dodge(0.9),angle=20)+ylim(c(0,23000))+
  scale_fill_manual(name="",values = c("Darkblue","green"),labels=c("Tiger","Leopard"))
```
Nearly 89% and 78% of the camera trap images of tigers and leopards are taken during night time, between 1800 and 0559 hrs. 


### 5. Distribution of Camera Trapped Images and Individuals among different sites
```{r diffrent sites, warning=F,echo=F,message=F,fig.width=10, fig.height=8}
## number of images 
require(dplyr)
require(ggplot2)
require(gridExtra)
camera_trap_data$name<-stringr::str_replace(camera_trap_data$name,"leopard","Leopard")
img_site<-group_by(camera_trap_data,site,name)%>%summarise(count=n(),disti_ind=n_distinct(id))%>%as.data.frame()
img_site<-na.omit(img_site)
img_site<-img_site[order(-img_site$count),]
img_site$site<-ordered(img_site$site,levels=unique(img_site$site))
levels(img_site$site)<-c("Nagarahole","Bandipur","Bhadra","Wayanad", "BiligiriRT", "Anshi-Dandeli","Kudremukh","CauveryMM",
                         "Tithimathi","Mhadei","Pench(Mah)", "Goa" )
img_site$name<-ordered(img_site$name,levels=c("Tiger","Leopard"))
plot_img<-ggplot(data = img_site,aes(x=site,y=count,fill=name))+geom_bar(stat = "identity",position = position_dodge(0.9))+theme_classic()+theme(axis.text.x = element_text(angle = 45,vjust = 0.5))+
  labs(x="Number of images",y="")+
  scale_fill_manual(name="",values=c("Darkblue","green"),labels=c("Tiger","Leopard"))
## number of individuals. 
plot_ind<-ggplot(data = img_site,aes(x=site,y=disti_ind,fill=name))+geom_bar(stat = "identity",position = position_dodge(0.9))+theme_classic()+theme(axis.text.x = element_text(angle = 45,vjust = 0.5))+
  labs(x="Number of individuals",y="")+
  scale_fill_manual(name="",values=c("Darkblue","green"),labels=c("Tiger","Leopard"))
grid.arrange(plot_img,plot_ind,nrow=1,ncol=2)

library(knitr)
names(img_site)<-c("Camera trapsites","Speceis","number of images","number of individuals")
kable(img_site[img_site[,2]=="Tiger",],row.names = F,caption = "Tiger images")
kable(img_site[img_site[,2]=="Leopard",],row.names = F,caption = "Leopard images")

```
Nearly 73% of the camera trapped images and 83% of the individual tigers, and, 60% of images and 56% of individual leopards are from Nagarahole, Bandipur and Wayanad sites.



